---
title: "Six Nations 2017 - Lineout Analysis"
author: 'Wal McConnell'
output:
  pdf_document: default
  html_document: default
date: "21 March 2017"
---

```{r setup, include=FALSE}
library(tidyverse)
library(rmarkdown)
library(knitr)
tournament_data = read.csv(file = '../../data/tournament.csv')
tournament_data <- tournament_data %>% separate(position, c('position_type', 'position_numnber'))
```
# Lineout

Ireland had a 91.7% success rate from the Lineout on their own ball, despite having the (joint) most lineouts (72) of the tournament. Tournament winners England lost only three of their own lineouts (2 v Ireland, 1 v France).

```{r, echo=FALSE}
lineouts_by_team <- tournament_data %>% 
  group_by(team) %>%
  summarise(
    total_lineouts = sum(total_lineouts),
    lineouts_won = sum(lineouts_won), 
    lineouts_lost = sum(lineouts_lost),
    win_percentage = round(lineouts_won/total_lineouts * 100, 1)
  ) %>% arrange(desc(win_percentage))

kable(lineouts_by_team, caption = "Lineouts By Team", col.names = c('Team', 'Total', 'Won', 'Lost', 'Win %'))
```

Irelands winning 22 from 22 lineouts v France was their best showing, while three of Irelands six lost lineouts occurred in their defeat by Wales 

```{r, fig.width=7, fig.height=4, echo = FALSE}
lineouts_won_by_opposition <- tournament_data %>% 
  group_by(team, opposition) %>%
  summarise(
    total_lineouts = sum(total_lineouts),
    lineouts_won = sum(lineouts_won), 
    lineouts_lost = sum(lineouts_lost),
    win_percentage = round(lineouts_won/total_lineouts * 100, 1)
  ) %>% arrange(desc(lineouts_lost), opposition)

ggplot(lineouts_won_by_opposition, aes(opposition, team)) + 
  geom_tile(aes(fill = win_percentage)) + 
  geom_text(aes(label = paste(lineouts_won, " / ", total_lineouts))) +
  scale_fill_gradient(low = "red", high = "green") +
  labs(fill='Win %') 
```

### Lineouts Stolen

Behind Italy, Ireland and Scotland jointly had the most lineouts stolen in the tournament

```{r, echo = FALSE}
lineouts_stolen_against <- tournament_data %>%
  group_by(opposition) %>%
  summarise(lineout_won_steal = sum(lineout_won_steal), stolen_percentage = round(((lineout_won_steal / sum(total_lineouts)) * 100), 1)) %>% 
  arrange(desc(lineout_won_steal)) %>% 
  filter(lineout_won_steal > 0)

kable(lineouts_stolen_against, caption = "Lineouts Stolen", col.names = c('Team', 'Lineouts Stolen', 'Stolen Lineout %'))
```


Ireland stole 4 opposition lineouts through Devin Toner, Iain Henderson and Peter O'Mahony. Englands Maro Itoje stealing 3 on the opposition throw (2 v Italy, 1 v Scotland).

```{r, echo = FALSE}
lineouts_stolen <- tournament_data %>%
  group_by(team, player_name) %>%
  summarise(lineout_won_steal = sum(lineout_won_steal)) %>% 
  arrange(desc(lineout_won_steal), team, player_name) %>% 
  filter(lineout_won_steal > 0)

kable(lineouts_stolen, caption = "Lineout Stolen", col.names = c('Team', 'Player', 'Lineouts Stolen'))
```

### Lineout Throwing Accuracy

```{r, echo = FALSE}
ireland_lineout_throwers <- tournament_data %>% 
  filter(team == 'Ireland') %>%
  group_by(player_name) %>%
  summarise(
    total_lineouts = sum(total_lineouts),
    lineouts_won = sum(lineouts_won), 
    lineouts_lost = sum(lineouts_lost),
    win_percentage = round(lineouts_won/total_lineouts * 100, 1)
  ) %>% 
  filter(total_lineouts > 0) %>% 
  arrange(desc(total_lineouts))
                
kable(ireland_lineout_throwers, caption = "Ireland - Lineout Throw Success", col.names = c('Player', 'Lineouts', 'Won', 'Lost', '% Success'))
```

Rory Best threw at 89% for the tournament, though the not straight lineout against England was intentional/tactical.

```{r, echo = FALSE}
rory_best_lineout_throw_lost <- tournament_data %>% 
  filter(team == 'Ireland', player_name == 'Best, Rory') %>%
  group_by(opposition) %>%
  summarise(
    total_lineouts = sum(total_lineouts),
    lineout_success = round(mean(lineout_success) * 100, 1),
    lineout_throw_lost_outright = sum(lineout_throw_lost_outright),
    lineout_throw_lost_handling_error = sum(lineout_throw_lost_handling_error),
    lineout_throw_lost_not_straight = sum(lineout_throw_lost_not_straight)
  ) %>% filter(lineout_success < 100)

kable(rory_best_lineout_throw_lost, caption = "Rory Best - Throw Success", col.names = c('Opposition', 'Total', '% Success', 'Lost Outright', 'Lost - Handling Error', 'Lost - Not Straight'))
```
